### 2017-09-22 Sections/physical_architectures_for_quantum_computing.tex
master text from Tim Byrnes - not edited in
indexing
Fusion strategies

\begin{figure}[!htb]
\includegraphics[width=0.9\columnwidth]{KLM_optical_paths}
\caption{Evolution of the four logical basis states through the KLM CZ gate. The NS gates do nothing in the first three cases, since they are operating only on vacuum and single-photon terms, which are left unchanged by the NS gate. In the last case, where both control and target are on, HOM interference results in photon bunching after the first beamsplitter, thereby creating two-photon terms. These terms inherit the $\pi$ phase-shift from the NS gate transformation, after which the final beamsplitter reverses the HOM photon bunching, yielding the same logical basis state with an acquired $\pi$ phase-shift.}. \label{fig:KLM_explain} \index{Knill-Laflamme-Milburn (KLM)}\index{CZ gate}
\end{figure}

@@

@@

It was then noted \cite{BrowneRudolph} that CZ gates aren't required at all for the preparation of optical cluster states. Instead, parity measurements (Sec.~\ref{sec:bell_proj}) operating in a rotated basis may be used to fuse smaller cluster states into larger ones, albeit acting destructively on two of the qubits, and also being non-deterministic, with a success probability of $1/2$. These gates have become known as \textit{fusion gates}, of which there are two types:
\begin{itemize}
	\item Type-I: destroy only a single photon, but require efficient number-resolved detection.
	\item Type-II: destroy two photons, but only require on/off detectors.
\end{itemize}
Both types of gates have several highly favourable characteristics:
\begin{itemize}
	\item Only HOM stability is required (Sec.~\ref{sec:opt_stab}). At no stage in the cluster state preparation procedure is any interferometric (i.e wavelength-scale) stability required.
	\item Gate failure is heralded by measurement of the wrong photon-number.
	\item Gate failure measures the respective qubits in the computational basis, thereby simply removing those qubits from the cluster state graph, whilst preserving the remainder of the state.
\end{itemize}

@@

Despite their non-determinism, numerous authors have examined approaches for efficiently preparing arbitrarily large cluster states using these destructive, non-deterministic gates.

@@

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{ 
function WhichPathErasure():
\begin{enumerate}
\item Alice and Bob each prepare an equal superposition of the two logical basis states,
\begin{align}
\ket\psi_\text{in} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1})\ket{0}_{B_2},
\end{align}
where $A_1/B_1$ denote the matter qubits, and $A_2/B_2$ denote their coupled optical modes.
\item Apply a $\pi$-pulse to each qubit, inducing a \mbox{$\ket{\!\downarrow}\to\ket{e}$} transition,
\begin{align}
\ket\psi_\pi = \hat{U}_\pi\ket\psi_\text{in} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{e}_{A_1})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{e}_{B_1})\ket{0}_{B_2}.
\end{align}
\item Wait for a coherent relaxation, inducing the transition \mbox{$\ket{e}\to\ket{\!\downarrow}\hat{a}^\dag$}, which emits a single photon,
\begin{align}
\ket\psi_\text{relax} = \hat{U}_\text{relax}\ket\psi_\pi = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1}\hat{a}^\dag_{A_2})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1}\hat{a}^\dag_{B_2})\ket{0}_{B_2}.
\end{align}
\item Apply a 50:50 beamsplitter between the two optical modes,
\begin{align}
\ket\psi_\text{BS} = \hat{U}_\text{BS} \ket\psi_\text{relax} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1}[\hat{a}^\dag_{A_2}+\hat{a}^\dag_{B_2}])\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1}[\hat{a}^\dag_{A_2}-\hat{a}^\dag_{B_2}])\nonumber \\
&\cdot \ket{0}_{A_2}\ket{0}_{B_2}.
\end{align}
\item Conditional upon detecting exactly one photon between the output optical modes, we obtain,
\begin{align}
\ket\psi_\text{out}^{1,0} = \bra{1,0}_{A_2,B_2} \ket\psi_\text{BS} = \frac{1}{2} (\ket{\!\uparrow,\downarrow}_{A_1,B_1} + \ket{\!\downarrow,\uparrow}_{A_1,B_1}), \nonumber \\
\ket\psi_\text{out}^{0,1} = \bra{0,1}_{A_2,B_2} \ket\psi_\text{BS} = \frac{1}{2} (\ket{\!\uparrow,\downarrow}_{A_1,B_1} - \ket{\!\downarrow,\uparrow}_{A_1,B_1}),
\end{align}
which is a Bell-pair between the matter qubits.
\end{enumerate}}}}
\caption{Using which-path erasure to entangle two $\lambda$-configuration matter qubits via post-selected linear optics. Note that the two matter qubits could in principle be arbitrarily physically separated. Only the emitted photons need be brought together locally for the implementation of a beamsplitter operation. This lends such entanglement generation protocols to distributed implementation.} \label{alg:which_path}
\end{table}

>>>>

\begin{figure}[!htb]
\includegraphics[width=0.9\columnwidth]{KLM_optical_paths}
\caption{Evolution of the four logical basis states through the KLM CZ gate. The NS gates do nothing in the first three cases, since they are operating only on vacuum and single-photon terms, which are left unchanged by the NS gate. In the last case, where both control and target are on, HOM interference results in photon bunching after the first beamsplitter, thereby creating two-photon terms. These terms inherit the $\pi$ phase-shift from the NS gate transformation, after which the final beamsplitter reverses the HOM photon bunching, yielding the same logical basis state with an acquired $\pi$ phase-shift.}. \label{fig:KLM_explain} \index{Knill-Laflamme-Milburn (KLM)}\index{Controlled-Z (CZ) gate}
\end{figure}

@@

\comment{To do! Talk about different gate types. Show update table. Reference from fusion strategies section.}

@@

It was then noted \cite{BrowneRudolph} that CZ gates aren't required at all for the preparation of optical cluster states. Instead, parity measurements (Sec.~\ref{sec:bell_proj}) operating in a rotated basis may be used to fuse smaller cluster states into larger ones, albeit acting destructively on two of the qubits, and also being non-deterministic, with a success probability of $1/2$. These gates have become known as \textit{fusion gates}, of which there are two types:
\begin{itemize}
	\item Type-I: destroy only a single photon, but require efficient number-resolved detection.
	\item Type-II: destroy two photons, but only require on/off detectors, since the gate succeeds upon coincidence events only.
\end{itemize}
Both types of gates have several highly favourable characteristics:
\begin{itemize}
	\item Only HOM stability is required (Sec.~\ref{sec:opt_stab}). At no stage in the cluster state preparation procedure is any interferometric (i.e wavelength-scale) stability required.
	\item Gate failure is heralded by measurement of the wrong photon-number.
	\item Gate failure measures the respective qubits in the computational basis, thereby simply removing those qubits from the cluster state graph, whilst preserving the remainder of the state, which can be `recycled'\index{Cluster state recycling} for reuse.
\end{itemize}

\comment{To do! Picture}

@@

Despite their non-determinism, numerous authors have examined approaches for efficiently preparing arbitrarily large cluster states using these destructive, non-deterministic gates \cite{???}. We will now briefly sketch out the formalism introduced by \cite{bib:RohdeBarrett07} for performing a theoretical analysis of such cluster state preparation schemes. This formalism was initially only presented in the context of linear clusters. However, these act as a primitive resource from which less trivial clusters may be constructed, such as rectangular lattices, which are universal for MBQC.

Consider a non-deterministic KLM-type CZ gate\footnote{In reality, no one would use KLM-type gates for preparing cluster states, owing to their complexity compared to fusion gates. Rather, we use this gate for illustrative purposes, since its operation upon success and failure are very simple for exposition.}, which upon failure destroys its two input photons by measuring them in the computational basis, and leaves the number of qubits unchanged upon success and bonds them together.

To analyze the operation of such a non-deterministic protocol, we begin by defining a vector $\vec{n}_t$ at time $t$, which stores purely classical information. Specifically, the vector tells us how many clusters of every length we have stored in memory (except single qubit clusters, which we assume `come for free'). For example, the second element of the vector tell us how many 3-qubit linear clusters we have in our possession.

We then define a strategy, $f_s$\index{Fusion strategies}, for choosing clusters we have stored and bonding them together to form larger clusters. Note that a key result from this study was that different strategies are hugely different in their ability to prepare long clusters. The strategy acts on our cluster vector and updates it accordingly,
\begin{align}\index{Fusion strategy update rules}
\vec{n}_{t+1} = f_s(\vec{n}_t).
\end{align}
That is, the length vector can be thought of as a series of `buckets' containing clusters of different lengths, and the strategy simply shifts the contents of the buckets around between buckets each time it is applied.

The strategy description, $f_s$, is also responsible for taking care of updating the elements of $\vec{n}$ according to an update rule, which dictates how many photons are lost or gained upon success or failure of the non-deterministic gate. For example, the CZ gate we have employed here for our toy model destroys two qubits upon failure, but upon success creates a cluster of length given by the sum of the lengths of the clusters acted upon by the gate.

We are then interested in the `rate'\index{Cluster state preparation rate}, $r$, at which large clusters are output from the protocol. This is simply extracted by defining a single parameter which counts the number of clusters in $\vec{n}$ above some predetermined target length, and normalises it by the total time taken to reach that point. The rate parameter converges asymptotically for long runtimes. Formally, the rate of preparation is given by,
\begin{align}
r = \lim_{t\to\infty} \frac{N_t}{t},
\end{align}
where $N_t$ is the total number of clusters of length greater than the target length at time $t$.

This completes the theoretical analysis for different strategies and gate types, allowing us to explore different approaches tailored to different physical systems and their varying gate implementations.

One of the key outcomes of this research was that a {\sc Balanced} strategy\index{Balanced strategy} is optimal in terms of preparation rate. This is simply a strategy which preferentially always bonds clusters of equal length.

@@

\begin{table}[!htb]
\fbox{\parbox{0.965\columnwidth}{\texttt{ 
function WhichPathErasure():
\begin{enumerate}
\item Alice and Bob each prepare an equal superposition of the two logical basis states,
\begin{align}
\ket\psi_\text{in} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1})\ket{0}_{B_2},
\end{align}
where $A_1/B_1$ denote the matter qubits, and $A_2/B_2$ denote their coupled optical modes.
\item Apply a $\pi$-pulse to each qubit, inducing a \mbox{$\ket{\!\downarrow}\to\ket{e}$} transition,
\begin{align}
\ket\psi_\pi = \hat{U}_\pi\ket\psi_\text{in} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{e}_{A_1})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{e}_{B_1})\ket{0}_{B_2}.
\end{align}
\item Wait for a coherent relaxation, inducing the transition \mbox{$\ket{e}\to\ket{\!\downarrow}\hat{a}^\dag$}, which emits a single photon,
\begin{align}
\ket\psi_\text{relax} = \hat{U}_\text{relax}\ket\psi_\pi = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1}\hat{a}^\dag_{A_2})\ket{0}_{A_2}\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1}\hat{a}^\dag_{B_2})\ket{0}_{B_2}.
\end{align}
\item Apply a 50:50 beamsplitter between the two optical modes,
\begin{align}
\ket\psi_\text{BS} = \hat{U}_\text{BS} \ket\psi_\text{relax} = &\frac{1}{2}(\ket{\!\uparrow}_{A_1}+\ket{\!\downarrow}_{A_1}[\hat{a}^\dag_{A_2}+\hat{a}^\dag_{B_2}])\nonumber \\
&\cdot (\ket{\!\uparrow}_{B_1}+\ket{\!\downarrow}_{B_1}[\hat{a}^\dag_{A_2}-\hat{a}^\dag_{B_2}])\nonumber \\
&\cdot \ket{0}_{A_2}\ket{0}_{B_2}.
\end{align}
\item Conditional upon detecting exactly one photon between the output optical modes, we obtain,
\begin{align}
\ket\psi_\text{out}^{1,0} = \bra{1,0}_{A_2,B_2} \ket\psi_\text{BS} = \frac{1}{2} (\ket{\!\uparrow,\downarrow}_{A_1,B_1} + \ket{\!\downarrow,\uparrow}_{A_1,B_1}), \nonumber \\
\ket\psi_\text{out}^{0,1} = \bra{0,1}_{A_2,B_2} \ket\psi_\text{BS} = \frac{1}{2} (\ket{\!\uparrow,\downarrow}_{A_1,B_1} - \ket{\!\downarrow,\uparrow}_{A_1,B_1}),
\end{align}
which is a Bell-pair between the matter qubits.
    \item $\Box$
\end{enumerate}}}}
\caption{Using which-path erasure to entangle two $\lambda$-configuration matter qubits via post-selected linear optics. Note that the two matter qubits could in principle be arbitrarily physically separated. Only the emitted photons need be brought together locally for the implementation of a beamsplitter operation. This lends such entanglement generation protocols to distributed implementation.} \label{alg:which_path}
\end{table}

### 2017-09-22 Sections/protocols_for_the_quantum_internet.tex
Fusion strategies

The cynics amongst us might question the non-determinism of the laws of Nature (`God does not play dice!' -- Albert Einstein), and ask whether quantum random numbers really are truly random (in the sense of non-determinism), or whether they also are just too hard to predict that we treat them as effectively random. The answer to this is that it has been proven that quantum mechanics is inconsistent with `hidden variable theories' \cite{Bell}, i.e that there is an underlying, but inaccessible determinism in the world, which is guiding quantum measurements in a completely deterministic manner. This disproof effectively validates the notion of quantum mechanical perfect random number generation.

@@

This scenario is an obvious example of where a UDP-like \textsc{Send And Forget} protocol may be viable. Unlike most other applications, Bob is broadcasting a stream of identical, pure quantum states, that are not entangled with any peripheral system, and are easily replicated, with no correlations between distinct photons. Thus, if any particular photon fails to reach Alice, it matters not, as she can simply await the next one emanating from Bob's bombardment of photons. There are no QoS requirements.

@@

Thus, in terms of Pauli errors, no special treatment is required by the QTCP protocol -- it is almost as if the teleportation protocol weren't there. And in terms of loss, the Bell state projection diagnoses lost qubits, allowing appropriate action to be taken, which is actually better than if the error were undiagnosed.

The total resources required to teleport a single-qubit state are:
\begin{enumerate}
\item The qubit to be teleported.
\item A shared Bell-pair.
\item A two-qubit measurement in the Bell basis.
\item The transmission of two classical bits.
\item Two classically-controlled Pauli gates.
\end{enumerate}

@@

Quantum key distribution (QKD) is a relatively mature technology with already several commercial systems being available off-the-shelf \cite{??? example companies}. The main drawback is thus the ability to perform long-distance transmission of photons. For these reasons, current QKD networks have been limited to relatively small distances, on the order of $\sim100$km, including in Austria, Switzerland, Japan, USA, and China \cite{bib:lo2014secure}. The longest distance QKD network that is currently planned is the Beijing to Shanghai QKD link spanning a distance of 2000km. This involves 32 trusted nodes\index{Trusted nodes} to break the total distance into shorter segments to convert the quantum information to classical information.

>>>>

The cynics amongst us might question the non-determinism\index{Non-determinism of quantum mechanics} of the laws of Nature (`God does not play dice!' -- Albert Einstein), and ask whether quantum random numbers really are truly random (in the sense of non-determinism), or whether they also are just too hard to predict that we treat them as effectively random. The answer to this is that it has been proven that quantum mechanics is inconsistent with `hidden variable theories' \cite{Bell}\index{Hidden variable theory}, i.e that there is an underlying, but inaccessible determinism in the world, which is guiding quantum measurements in a completely deterministic manner. This disproof effectively validates the notion of quantum mechanical perfect random number generation.

@@

This scenario is an obvious example of where a UDP-like \textsc{Send-and-forget} protocol may be viable. Unlike most other applications, Bob is broadcasting a stream of identical, pure quantum states, that are not entangled with any peripheral system, and are easily replicated, with no correlations between distinct photons. Thus, if any particular photon fails to reach Alice, it matters not, as she can simply await the next one emanating from Bob's bombardment of photons (the `shotgun' approach\index{Shotgun protocol}). There are no QoS requirements.

@@

Thus, in terms of Pauli errors, no special treatment is required by the QTCP protocol -- it is almost as if the teleportation protocol weren't there. And in terms of loss, the Bell state projection diagnoses lost qubits, allowing appropriate action to be taken, which is actually better than if the error were undiagnosed. These are often referred as \textit{located}\index{Located errors} and \textit{unlocated}\index{Unlocated errors} errors.

The total resources required to teleport a single-qubit state are:
\begin{enumerate}
\item The qubit to be teleported.
\item A shared Bell-pair.
\item A two-qubit entangling measurement in the Bell basis.
\item The transmission of two classical bits.
\item Two classically-controlled Pauli gates for correction.
\end{enumerate}

@@

Quantum key distribution (QKD) is a relatively mature technology with already several commercial systems being available off-the-shelf \footnote{\cite{??? example companies}}. The main drawback is thus the ability to perform long-distance transmission of photons. For these reasons, current QKD networks have been limited to relatively small distances, on the order of $\sim100$km, including in Austria, Switzerland, Japan, USA, and China \cite{bib:lo2014secure}. The longest distance QKD network that is currently planned is the Beijing to Shanghai QKD link spanning a distance of 2000km. This involves 32 trusted nodes\index{Trusted nodes} to break the total distance into shorter segments to convert the quantum information to classical information.

### 2017-09-22 Sections/the_quantum_space_race.tex
Fusion strategies

In general, the distance to the horizon $d$ (in kilometres)\index{Distance to horizon} on the Earth's surface at an altitude of $h$ (in metres) is given by the approximation,
\begin{align}
d \approx 3.57\sqrt{h}.
\end{align}
This relationship is shown in Fig.~\ref{fig:dist_hor}, covering the range of altitudes from sea level to the summit of Mount Everest (8,848m). It is evident that line-of-sight channels across the Earth's surface are inherently limited by the curvature of the Earth to $\sim$250km, using altitudes that are practically realistic (i.e no ground stations on the summit of Everest!).

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{distance_to_horizon}
\caption{Distance to the horizon $d$ (in kilometers), at altitude $h$ (in meters). The range for $h$ spans from sea level to roughtly the height of Mount Everest, thereby covering the full range of altitudes possible for ground-based network nodes. However, using more practically realisitic bounds on altitude, say on the order of 5,000-6,000m (already very optimistic!), line-of-sight distance to the horizon will be limited to be on the order of 200-250km. This limitation necessitates the use of repeater networks to build up quantum communication channels that bypass this constraint imposed by the Earth's curvature.} \label{fig:dist_hor}	
\end{figure}

>>>>

In general, the distance to the horizon $d$ (in kilometres)\index{Distance to horizon} on the Earth's surface at an altitude of $h$\index{Ground station altitude} (in metres) is given by the approximation,
\begin{align}
d \approx 3.57\sqrt{h}.
\end{align}
This relationship is shown in Fig.~\ref{fig:dist_hor}, covering the range of altitudes from sea level to the summit of Mount Everest (8,848m). It is evident that line-of-sight channels across the Earth's surface are inherently limited by the curvature of the Earth to $\sim$250km, using altitudes that are practically realistic (i.e no ground stations on the summit of Everest!).

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{distance_to_horizon}
\caption{Distance to the horizon $d$ (in kilometers), at altitude $h$ (in meters), any several illustrative examples with their elevation. The range for $h$ spans from sea level to roughtly the height of Mount Everest, thereby covering the full range of altitudes possible for ground-based network nodes. However, using more practically realisitic bounds on altitude, say on the order of 5,000-6,000m (already very optimistic!), line-of-sight distance to the horizon will be limited to be on the order of 200-250km. This limitation necessitates the use of repeater networks to enable the efficient approaches for constructing quantum communication channels that bypass this constraint imposed by the Earth's curvature.} \label{fig:dist_hor}
\end{figure}
